A thesis draft

### 2018-11-18 res/sections/08-Background.tex
Add part on network slicing

@@

\subsection{SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be split in three different layers, called planes, based on 
their functionalities.

@@

The \emph{Application plane} role is to set the behaviour of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way.
Analyzing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previously defined NBIs.
These applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.

>>>>

\section{Network slicing}
A network slice can be defined as a logical virtual network, tailored for a
given use case, on the top a common network using technologies such as
Software-Defined Networking and Network Function Chaining that allows network
softwarization~\cite{ordonez2017network}. The notion of separated virtual
networks deployed on the same infrastructure is not new, e.g. VPN, but it has
some specifics that make it a novel idea. The concept of network slicing was
first introduced in 2002 with the paper~\cite{peterson2003blueprint}, in which
is described an overlay network that allow services to access different slices
of overlay's resources and in which overlay management services run in their
own slice. This idea was be refined during time and
in~\cite{alliance2016description} the design of network slicing is based on a
three-layers model, as depicted in
Figure~\ref{chap:background:img:network_slicing}.

\begin{figure}
  \centering
  \includegraphics[scale=0.2]{networkslicing}
  \caption{The three layers model of network slicing}
  \label{chap:background:img:network_slicing}
\end{figure}

The first layer is the \emph{Service Instance Layer} used to represent the
business or user services supported, the second one is the \emph{Network Slice
Instance Layer}. This is a set of network functions that allows to
satisfy certain requirements~\cite{kotulski2017end}. Network functions, in
general, can be both physical, thus a combination of vendor-specific hardware
and/or virtualized. In that case the function is implemented as software and it
is decoupled from the hardware on which it is running on. Each instance must be
isolated, in order to ensure that function running at the top of it will always
provide the same Quality of Service, without degrading because of a change on
another slice~\cite{richart2016resource}. It also allows to provide better security. In fact, in case of attack or fault of a certain slice, others must not be affected. Moreover, each slice must implement it own set of security function to preventing unauthorized access. Furthermore, isolation let to manage individually each slice~\cite{ordonez2017network}. Network slice instances are defined by a \emph{Network Slice Blueprint}: a complete description of the structure and data flows that allow the creation and the control of the Network Slice Instance during its whole lifecycle. The blueprint also refers to the resources needed to assemble the network slice and defines its characteristics (e.g. ultra-low latency, ultra-reliability, value-added services for enterprises, etc.). The last layer is the \emph{Resource Layer}, containing both physical and logical resources, that involve all physical assets and resources for computation, storage or transport, including radio access.

@@

\subsection{SDN stack}
\begin{figure}[ht]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be split in three different layers, called planes, based on 
their functionalities.

@@

The \emph{Application plane} role is to set the behavior of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way.
Analyzing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previously defined NBIs.
These applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.
